Notes on Control Plane Inference

### 2021-03-29 inference.tex
add p4v discussion (sans proof)

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}

@@

\begin{figure}
  \[\begin{array}{r>{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Prog \to \GCL} \\
  (h.f := e)[\sigma] && h.f := e   \\
  (\assume b)[\sigma] && \assume b \\
  (\assert b)[\sigma] && \assert b \\
  (c_1;c_2)[\sigma] && c_1[\sigma];c_2[\sigma] \\
  (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
  (\choose \rho t c)[\sigma] && \textrm{\textcolor{red}{unnecessary, but similar to t.apply}}\\
  (t.\apply())[\sigma] &&
  \begin{array}[t]{l}
    \{\\
    \quad \assume {r_1.\matches = t.\keys};\\
    \quad t.\action[r_1.\action](r_1.\data)\\
    \}~\choice\\
    \multicolumn{1}{c}{\vdots}\\
    \choice~\{\\
    \quad \assume {r_n.\matches = t.\keys};\\
    \quad t.\action[r_n.\action](r_n.\data)\\
    \}\\
    \textrm{where}~[r_1;\ldots;r_n] = \sigma(t.\id)
  \end{array}
  \end{array}
  \]
  \caption{Table Elimination}
  \label{fig:table-elim}
\end{figure}

@@

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{(\cdot)[?] : \Prog \to \Instr} \\
  (h.f := e)[?] && h.f := e \\
  (\assume b)[?] && \assume b \\
  (\assert b)[?] && \assert b\\
  (c_1;c_2)[?] && c_1[?];c_2[?]\\
  (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
  (\choose \rho t c)[?] &&
  \textrm{\textcolor{red}{unnecessary}} \\
  (t.\apply())[?] &&
  \begin{array}[t]{l} \displaystyle
    \{ \\
    \quad \choose \rho t \;\{ \\
    \qquad \assume \rho.\matches = t.\keys; \\
    \qquad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
    \left\{\begin{array}{l}
    \assume \rho.\action = i;\\
    t.\action[i](\rho.\data)
    \end{array}\right\}\\
    \quad \} \\
    \} \choice \{ \\
    \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
    \quad \SKIP \\
    \quad \}\\
    \}
  \end{array}
  \end{array}
  \]
  \caption{Symbolic Table instrumentation}
  \label{fig:table-instrument}
\end{figure}

@@

>>>>

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
\newcommand{\pv}{\mathit{p4v}}

@@

\begin{figure}
  \[\begin{array}{r>{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Prog \to \GCL} \\
  (h.f := e)[\sigma] && h.f := e   \\
  (\assume b)[\sigma] && \assume b \\
  (\assert b)[\sigma] && \assert b \\
  (c_1;c_2)[\sigma] && c_1[\sigma];c_2[\sigma] \\
  (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
  (\choose \rho t c)[\sigma] && \textrm{\textcolor{red}{undefined}}\\
  (t.\apply())[\sigma] &&
  \begin{array}[t]{l}
    \{\\
    \quad \assume {r_1.\matches = t.\keys};\\
    \quad t.\action[r_1.\action](r_1.\data)\\
    \}~\choice\\
    \multicolumn{1}{c}{\vdots}\\
    \choice~\{\\
    \quad \assume {r_n.\matches = t.\keys};\\
    \quad t.\action[r_n.\action](r_n.\data)\\
    \}\\
    \textrm{where}~[r_1;\ldots;r_n] = \sigma(t.\id)
  \end{array}
  \end{array}
  \]
  \caption{Table Elimination}
  \label{fig:table-elim}
\end{figure}

@@

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{(\cdot)[?] : \Prog \to \Instr} \\
  (h.f := e)[?] && h.f := e \\
  (\assume b)[?] && \assume b \\
  (\assert b)[?] && \assert b\\
  (c_1;c_2)[?] && c_1[?];c_2[?]\\
  (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
  (\choose \rho t c)[?] &&
  \textrm{\textcolor{red}{undefined}} \\
  (t.\apply())[?] &&
  \begin{array}[t]{l} \displaystyle
    \{ \\
    \quad \choose \rho t \;\{ \\
    \qquad \assume \rho.\matches = t.\keys; \\
    \qquad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
    \left\{\begin{array}{l}
    \assume \rho.\action = i;\\
    t.\action[i](\rho.\data)
    \end{array}\right\}\\
    \quad \} \\
    \} \choice \{ \\
    \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
    \quad \SKIP \\
    \quad \}\\
    \}
  \end{array}
  \end{array}
  \]
  \caption{Symbolic Table instrumentation}
  \label{fig:table-instrument}
\end{figure}

@@

\section{P4v comparison}

Ive said before that the P4v-style instrumentation is an overapproximation of
the control plane behavior. It's a slightly inaccurate claim in that it
correctly captures all possible control plane behaviors, but it doesn't have a
handle on constrain control plane behavior.

A simplified version of P4v's encoding is shown in Figure~\ref{fig:p4vlike}. It
omits the key-reads commands and hit/miss predicates.

\begin{figure}
  \[\begin{array}{l >{\triangleq}cl}
  \multicolumn{3}{l}{\pv : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{\pv : \Prog \to \Instr} \\
  \pv(h.f := e) && h.f := e \\
  \pv(\assume b) && \assume b \\
  \pv(\assert b) && \assert b \\
  \pv(c_1;c_2) && \pv(c_1;c_2) \\
  \pv(c_1 \choice c_2) && \pv(c_1) \choice \pv(c_2) \\
  \pv(\choose \rho t c) && \textcolor{red}{\text{undefined}} \\
  \pv(t.\apply()) && \\
  \multicolumn{3}{>{\qquad}r}{
    \displaystyle
    \bigchoice_{i = 0}^{|t.\action|-1}t.\action[i]()
  }
  \end{array}\]
  \caption{A \texttt{p4v}-like instrumentation. For simplicity of reasoning
    about variables, assume that actions take no action data. }
  \label{fig:p4vlike}
\end{figure}

The theorem below demonstrates the over-approximation. Compare with the Instrumentation Equivalence Theorem~\ref{thm:instr-equiv}.

\begin{theorem}[\texttt{p4v} over-approximates]
  \label{thm:p4v-over}
  For every $c \in \Prog$, well-formed state $\sigma : \Value \to \Row^* $,
  model $\mu \in \Model$, and packet $\pkt \in \Pkt$,
  $\denote{c}^\sigma\,\mu\,\pkt \subseteq \denote{\pv(c)}^\sigma\,\mu\,\pkt$
\end{theorem}

\begin{proof}
\todo[inline]{By induction on c}.
\end{proof}

We can witness non-equality by the following counter-example. Consider a program
$t.\apply()$ where $t = \{\keys=[h.f];\actions=[\lambda().\SKIP, \lambda(). h.f :=
  99]\}$. Consider an input packet with $\pkt(h.f) = 47$, and an empty input
model and state that selects a single action.

Then $\denote{t.\apply()}^\emptyset\,\emptyset\,\pkt = \{(\mu,\pkt)\}$.

Whereas
\[\begin{array}{l}
\denote{\pv(t.\apply())}^\emptyset\,\emptyset\,\pkt \\
= \denote{
    \displaystyle
    \bigchoice_{i = 0}^{|t.\action|-1} t.\action[i]()
}^\emptyset\,\emptyset\,\pkt \\
= \denote{\SKIP}^\emptyset\,\emptyset\,\pkt \cup \denote{h.f := 99}^\emptyset\,\emptyset\,\pkt\\
= \{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}
\end{array}\]

Finally, $\{(\emptyset,\pkt)\} \subset
\{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}$, exemplifies a strict
subset behavior described by Theorem~\ref{thm:p4v-over}.
